---
title: "Houston Drivers Rough"
subtitle: "A Bayesian Analysis of Moving Violations in Houston"
author: "Will Curkan"
date: "2023-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The "greater Houston area" is a 665 $mi^2$ section of land surrounded and surrounding many suburban municipalities. TALK ABOUT ITS POPULATION AND THE SURROUNDING SUBURBS POP'NS.

## Introduction


Houston is an enormous metropolitan completely surrounded by suburban areas, denoted "the greater Houston area". The population and the automobile traffic it generates, getting from a suburban area on the south side to one on the north side can consume, more-or-less, an hour and a half of time. With this said, there are lots of cars which means lots of people rushing to get from one municipality or suburb to another, or to go in and out of Houston proper. The want to get from point A to point B in the 665 square mile city causes a subconscious anxiety and desire to rush, which causes police to give many speeding violations. Conveniently, police frequently record the speed of the offender and the posted speed limit when writing a citation, and this data is available via the Stanford Open Policing Project, which the Houston police department participates with in open-sourcing their data.

## Problem

Size and roads cause tons of traffic, accidents. Size is so big it takes forever to get places

## Purpose

See if a Bayesian analysis works on the data

```{r libraries}
library(dplyr, quietly = T)
```

```{r data}
data <- read.csv('tx_houston_2023_01_26.csv')
```

