---
title: "San Antonio drivers"
author: "Will Curkan"
date: "2023-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr, quietly = T)
```

1. Use standard updating to go through the years and take the average number of stops per year.

Only look at type because problem is pretty open otherwise. Look at Traffic violations. Specifically speeding?

600k observations after filtering NAs

Questions about speeding in HOUSTON?
- Mean ratio of speed to posted speed to negate the differences between high/low speed limits
  - Normal normal
- Mean time between speeding stops
  - GammaExp
- Number per day
  - GammaPoisson
- Speeding among other traffic related offenses
  - Beta-bin (Also poisson?) Check which model is better with this one
- Use predictive distributions.

Methods:

Use jags and std simulation for everything


Remove the features `raw_row_number`, `location`, `lat`, `lng`, `geocode_source`. (or just write the ones we will keep)

```{r}
data <- read.csv('tx_houston_2023_01_26.csv')
```

```{r}
colnames(data)
```

```{r}
data %>%
  group_by(type) %>%
  summarize(num_stops = n_distinct(time))
```

```{r}
data %>%
  group_by(violation) %>%
  summarize(n_vio = n_distinct(violation))
```

```{r rawSpeed_postedSpeed_raio}
# ratio of raw speed to posted speed limit
raw_posted_ratio <- data$speed / data$posted_speed

# Remove the speed ratios that became NA
# due to an undocumented speed (could be many reasons like it not being a speeding violation, or officer error)
raw_posted_ratio <- raw_posted_ratio[!is.na(raw_posted_ratio)]

raw_posted_ratio[1:5]
length(raw_posted_ratio)
```

```{r point_estimate_mean_ratio}
mean(raw_posted_ratio)
# People stopped are going approximately 34% higher than the speed limit
```

```{r}
# c <- data %>%
#   filter(type == 'vehicular') %>%
#   select(speed, posted_speed) %>%
#   filter(!is.na(speed))

data

# d <- c$speed - c$posted_speed
# sum(d != 0)
```

```{r}
NA == TRUE
```

```{r}
# Features to keep
keep <- c('date', 'time', 'subject_age', 'subject_race', 'subject_sex',
          'type', 'violation', 'arrest_made', 'citation_issued', )
# Subset dataframe with only features to keep
df <- data %>%
  select()
```

```{r}
v <- data %>%
  group_by(date) %>%
  summarize(num_stops = n_distinct(time))
```